### 2019-05-16 src/diagrams/domjudge-queue.tex
move image and table implementation into other files

\begin{tikzpicture}
    \begin{axis}[xlabel={Waktu (detik)}, ylabel={Jumlah jawaban pada antrian}, xmin=0,xmax=700, ymin=0, ymax=100, legend pos=north east, ymajorgrids=true, grid style=dashed,]
        \addplot[color=blue,] table {data/domjudge-queue-test1.txt};
        \addlegendentry{Pengujian pertama}
        \addplot[color=red,] table {data/domjudge-queue-test2.txt};
        \addlegendentry{Pengujian kedua}
        \addplot[color=green] table {data/domjudge-queue-test3.txt};
        \addlegendentry{Pengujian ketiga}
    \end{axis}
\end{tikzpicture}