### 2019-06-19 src/diagrams/ugrade-domjudge-queue.tex
add overlay diagram between ugrade queue and domjudge queue

\begin{tikzpicture}
    \begin{axis}[xlabel={Waktu (detik)}, ylabel={Jumlah jawaban pada antrian}, xmin=0,xmax=625, ymin=0, ymax=80, legend pos=north east, ymajorgrids=true, grid style=dashed,height=0.5\textwidth,width=\textwidth]
        \addplot[color=blue,] table {data/domjudge-queue-test2.txt};
        \addlegendentry{DOMJudge}
        \addplot[color=red,] table {data/ugrade-queue-gz5-test2.txt};
        \addlegendentry{UGrade dengan \textit{grading size} 5}
    \end{axis}
\end{tikzpicture}